---
title: "Data Exploration"
author: "Zoe Schroder"
date: "4/4/2019"
output: html_document
---

This .Rmd file is used to explore my data further. I hope to find some interesting results within my data that inspire new research questions. 

I am using `BigDays.RData` from the pred-casualties repository. Each row in the data frame **BigDays.sfdfT** represents a big tornado day defined as having 10 or more tornadoes on a given convective day (6 AM to 6 AM). Included are the maximum and average values of the atmospheric variables within the hull/ spatial exent of the event. (Convective Available Potential Energy (CAPE), Convective Inhibition (CIN), Helicity (HLCY), U-direction of storm motion (USTM), V-direction of storm motion (VSTM)...calculated by me are Bulk Shear (BS) and Storm Motion (SM)). The **Torn.sfT** data frame has information on every tornado that falls within a big day. This is point data...start of the tornado's genesis location.

Other variables to consider including: Lifted Condesation Levels (LCL), Significant Tornado Parameter (STP), Supercell Composite Parameter (SCP). Additionally varying the parcel level may prove interesting (most unstable, surface-based, mixed layer). 

To start any analysis, load the Big Day data. 
```{r}
load("BigDays.RData")
dim(BigDays.sfdfT)
dim(Torn.sfT)
```

Plot histograms of the data
```{r}

```

################
## Questions: ##
################

Is there a monthly variation in the atmospheric environments? 
November has a similar ATP to May. Are the environments similar in these months? 

What are the land atmospheric interactions that could influence the power on these big days? 

Is the duration of the event influenced by the environments that are present? 

Can you use the atmospheric environments to predict the number of tornadoes that will occur? 

Does the starting hour of the event influence the duration of the event? 

Is the area of the event correlated with the duration of the event? 