---
title: "Data Exploration"
author: "Zoe Schroder"
date: "4/4/2019"
output: html_document
---

This .Rmd file is used to explore my data further. I hope to find some interesting results within my data that inspire new research questions. 

I am using `BigDays.RData` from the pred-casualties repository. Each row in the data frame **BigDays.sfdfT** represents a big tornado day defined as having 10 or more tornadoes on a given convective day (6 AM to 6 AM). Included are the maximum and average values of the atmospheric variables within the hull/ spatial exent of the event. (Convective Available Potential Energy (CAPE), Convective Inhibition (CIN), Helicity (HLCY), U-direction of storm motion (USTM), V-direction of storm motion (VSTM)...calculated by me are Bulk Shear (BS) and Storm Motion (SM)). The **Torn.sfT** data frame has information on every tornado that falls within a big day. This is point data...start of the tornado's genesis location.

Other variables to consider including: Lifted Condesation Levels (LCL), Significant Tornado Parameter (STP), Supercell Composite Parameter (SCP). Additionally varying the parcel level may prove interesting (most unstable, surface-based, mixed layer). 

To start any analysis, load the Big Day data. 
```{r}
load("BigDays.RData")
dim(BigDays.sfdfT)
dim(Torn.sfT)
```

Plot histograms of the data
```{r}
BigDays.sfdfT$Duration <- as.numeric(BigDays.sfdfT$Duration)
hist(BigDays.sfdfT$Duration)
hist(BigDays.sfdfT$casualties)
hist(BigDays.sfdfT$nT)
hist(BigDays.sfdfT$maxEF)
```

################
## Questions: ##
################

Is there a monthly variation in the atmospheric environments? 
November has a similar ATP to May. Are the environments similar in these months? 

```{r}
library(dplyr)
month_table <- BigDays.sfdfT %>%
  group_by(Month) %>% 
  summarize(avgcape <- mean(maxCAPE),
            avgcin <- mean(minCIN),
            avghlcy <- mean(maxHLCY),
            avgbs <- mean(avgBS),
            avgsm <- mean(avgSM))
```

Is there a spatial dependency on the atmospheric variables?

```{r}
state_table <- Torn.sfT %>%
  group_by(st) %>% 
  summarize(avgED<- mean(ED) / 10^9,
            avgpathwidth <- mean(Length),
            avgcas <- mean(cas),
            avghour <- mean(Hour))

state_table 
```

Pull ED for each state from the Torn.sfT annd plot the average of the max ratings for each state. 

```{r}

```

Plot the total number of each EF ratings by state. MS had XX EF 0, MS had XX EF 1, MS had XX EF2, etc.

```{r}

```

What are the land atmospheric interactions that could influence the power on these big days? 

```{r}

```

Is the duration of the event influenced by the environments that are present? 

```{r}

```

Can you use the atmospheric environments to predict the number of tornadoes that will occur? 

```{r}

```

Does the starting hour of the event influence the duration of the event? 

```{r}

```

Is the area of the event correlated with the duration of the event? 

```{r}

```

